Type-Safe Modular Parsing

### 2017-01-14 paper/sections/RelatedWork.tex
Fix

\paragraph{Syntactically Extensible Parsing} Extensible parsing is
used in many different areas. Extensible parser generators are a
mainstream area specially designed for modular syntax and
parsing. Many parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
modular grammars. They allow users to create new modules where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars.  NOA~\cite{Gouseti2014} uses Java
annotation processing to gather information about grammars in Java
code. After all information is collected NOA merges all grammar
components, produce an ANTLR~\cite{antlr1995} grammar, and generates
corresponding parsing code from the ANTLR grammar.  Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking.

Macro systems like the C preprocessor, C++ templates and
Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming
techniques are a similar area aiming at syntactic extensibility.
SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar for
Java using library imports. Composition of syntactic sugar is easy for
users, but it requires many rounds of parsing and adaption. This
significantly affects efficiency of compilation. Since the
implementation was based on SDF~\cite{Heering1989} and
Stratego~\cite{Visser2001}, the approach does not support separate
compilation. Racket adopts a macro system for library-based language
extensibility~\cite{Tobin-Hochstadt2011}. This is quite powerful and it uses
attributed ASTs for contextual
information. Moreover extensions can be integrated in a modular
way. However such modularity is not flexible enough for language
unification, as the syntax is only built from extensions.
%Moreover,
%existing macros cannot be further changed after
%definition. \haoyuan{correct?}
Extensible
compilers like JastAdd~\cite{Ekman2007} and
Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
is mostly done using parser generators. They focus on the
extensions to a host language. Our technique is designed for type-safe
modular language parsers (not just language
extensions), with very flexible parser composition.

@@

\paragraph{Extensible Parsing Algorithms}
Extensible parsing algorithms are another area that relates to
separate compilation.
\textit{Parse table composition}~\cite{bravenboer2008parse}
is an approach where grammars are compiled to
modular parse tables. Those parse tables are expressed as DFAs
or NFAs, and later they can be composed by an algorithm, to provide
separate compilation for parsing. The generation of parse tables can
be quite expensive in terms of performance. The approach
is quite different from ours, since it uses parse
 tables, whereas we use parser combinators.
Our approach supports both
\emph{separate compilation} as well as \emph{modular
  type-checking}. Moreover, the extensibility of parsing is further
available at language composition and lexical level.

>>>>

\paragraph{Syntactically Extensible Parsing} Extensible parsing is
used in many different areas. Extensible parser generator is a
mainstream area specially designed for modular syntax and
parsing. Many parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016} support
modular grammars. They allow users to create new modules where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars.  NOA~\cite{Gouseti2014} uses Java
annotation processing to gather information about grammars in Java
code. After all information is collected, NOA merges all grammar
components, produce an ANTLR~\cite{antlr1995} grammar, and generates
corresponding parsing code from the ANTLR grammar. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking.

Macro systems like the C preprocessor, C++ templates and
Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming
techniques are a similar area aiming at syntactic extensibility.
SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar for
Java using library imports. Composition of syntactic sugar is easy for
users, but it requires many rounds of parsing and adaption. This
significantly affects efficiency of compilation. Since the
implementation was based on SDF~\cite{Heering1989} and
Stratego~\cite{Visser2001}, it does not support separate
compilation. Racket adopts a macro system for library-based language
extensibility~\cite{Tobin-Hochstadt2011}. This is quite powerful and it uses
attributed ASTs for contextual
information. Moreover extensions can be integrated in a modular
way. However such modularity is not flexible enough for language
unification, as the syntax is only built from extensions.
%Moreover,
%existing macros cannot be further changed after
%definition. \haoyuan{correct?}
Extensible
compilers like JastAdd~\cite{Ekman2007} and
Polyglot~\cite{Nystrom2003} also support extensible parsing, but this
is mostly done using parser generators. They focus on the
extensions to a host language. Our technique is designed for type-safe
modular language parsers (not just language
extensions), with very flexible parser composition.

@@

\paragraph{Extensible Parsing Algorithm}
Extensible parsing algorithm is another area that relates to
separate compilation.
\textit{Parse table composition}~\cite{bravenboer2008parse}
is an approach where grammars are compiled to
modular parse tables. Those parse tables are expressed as DFAs
or NFAs, and later they can be composed by an algorithm, to provide
separate compilation for parsing. The generation of parse tables can
be quite expensive in terms of performance. The approach
is quite different from ours, since it uses parse
 tables, whereas we use parser combinators.
Our approach supports both
\emph{separate compilation} as well as \emph{modular
  type-checking}. Moreover, the extensibility of parsing is further
available at language composition and lexical level.
